---
title: "Geocoding with R"
subtitle: "The SF package"
author: "Andrew Wells & Fernanda Candido Gomes"
institute: "Hertie School | [I2DS Tools for Data Science workshop](https://github.com/intro-to-data-science-21-workshop)"
date: "04th November 2021"
output: 
  html_document:
    toc: TRUE
    number_sections: FALSE
    highlight: tango
    theme: lumen
    toc_depth: 2
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
library(pacman)
pacman::p_load(sf, leaflets, readxl,
               dplyr, qtm, ggplot2, rvest, tidyverse, raster, tmap, tmaptools)
#raster is for mac

options(scipen = 999)
```

## Finding coordinates from addresses
As Fernanda said, geocoding is about taking an address or a name of a place and turning it into a geographic position on the earth’s surface. This means that if you run an address, a location, or landmark through the proper function in R, it will give us the coordinates, which can be mapped.

For this tutorial, we are focusing on mapping institutions that handle refugee policy and offices that process asylum claims. Below are the addresses of some of these institutions:

UNHCR Germany - Zimmerstrasse 80, Berlin, Germany
UNHCR Germany/Federal Office for Migrantion and Refugees - Frankenstrasse 210, Nurnberg, Germany
French Office of Immigration and Integration - 83 Rue de Patay, Paris, France
French Office of Immigration and Integration - 4 Rue Gustave Doré, Strasbourg, France
Belgium Immigration Office - Pachéco 44, Brussels, Belgium
Port of Antwerp Schengen Immigration Office - Kauwenstein 8, Antwerpen, Belgium
CARE - Switzerland - Chemin de Balexert 7, Geneva, Switzerland
CARE UK - 89 Albert Embankment, London, United Kingdom
Danish Refugee Council - Borgergade 10, København, Denmark

If we don't have the address, we can simply put the location, such as a city:
European Asylum Support Office - Valletta, Malta
Rome, Italy

Or we can put a landmark, such as:
La Tour Eiffel (Eiffel Tower)
Kościół Mariacki (Mary's Basilica in Krakow)
წმინდა სამება (Gergeti Trinity Church)

We simply same all of these address/locations as a single vector. We then run this vector through the
geocode_OSM() function. This function does the magic of geo-locating all of the information in the vector.

The output will be a data frame with the longitudes and latitudes of the locations. 

```{r pressure, echo=FALSE}
locations <- c("Zimmerstrasse 80, Berlin, Germany",
               "Frankenstrasse 210, Nurnberg, Germany",
               "83 Rue de Patay, Paris, France",
               "4 Rue Gustave Doré, Strasbourg, France",
               "Pachéco 44, Brussels, Belgium",
               "Kauwenstein 8, Antwerpen, Belgium",
               "Chemin de Balexert 7, Geneva, Switzerland",
               "89 Albert Embankment, London, United Kingdom",
               "Borgergade 10, København, Denmark",
               "Valletta, Malta",
               "Rome, Italy",
               "La Tour Eiffel",
               "Kościół Mariacki",
               "წმინდა სამება") 

geocoded_locations <- geocode_OSM(locations) #geocode_OSM() from tmap package

geocoded_locations
```

## Setting up the geo-points

We then use the st_as_sf() function to set the coordinates of the longitude and latitude as relative points on the x and y axes. 
This gives us the points with the critical "geometry" column, which can be used to plot the points on a map, which has a shapefile. 
It is also good to set the coordinated reference system (CRS) code within this function to 4326, which is standard for the World Geodetic System.

```{r pressure, echo=FALSE}
point_geo <- st_as_sf(geocoded_locations,
                      coords = c(x = "lon", y = "lat"),
                      crs = 4326) 

point_geo
```

## Downloading a map of the world
diva-gis.org is great for individual countries. But does not have easily accessible .shp files for more than one country. For this, let's go to Natural Earth (naturalearthdata.com). 

Click on the 1:50m Cultural Vectors. Right click on the "Download countries" button and copy the link. We will then follow the same processes as in the country .shp file example above to get geographical information for thw whole world.

```{r pressure, echo=FALSE}

world_file <- "" 

# Download .zip
download.file(world_file, destfile = "WORLD.zip")

# Unzip the file
unzip("WORLD.zip")

# Examine the file .zip file to find the shapefile
list.files()

# Save the shapefile as a data frame using the read_sf() function
world <- sf::read_sf("") #Look for the .shp
world
```


## Make a quick map of the world
Just to inspect that the data we downloaded is correct/functional, let's plot it quickly. We can do this in a number of different ways, below at three alternative ways: two with ggplot() and one with qtm(), which comes in the tmap package and makes a "quick thematic map".

```{r pressure, echo=FALSE}
ggplot(world) + geom_sf() + theme_minimal()

#Or
ggplot() + geom_sf(data = world) + coord_sf(crs = st_crs(4326)) + theme_minimal()

#Or
qtm(world) #qtm() function "Quick Thematic Map" from the tmap package allows us to quickly examine a map from the imported shapefile
```


## Joining dataframes

The world .shp file that we downloaded conveniently also has a "geometry" column, similar to that which we created using the st_as_sf function when creating the point_geo data frame.

Now that we have the point_geo and world data and they both have the "geometry" column, we can join then using the st_join() function, which comes in the sf package. st_join allows us to make a left or inner join with spatial data.

```{r pressure, echo=FALSE}
st_join(point_geo, world, join = st_within)
```

## Map it all together!

Using the tm_shape() function, from the tmap package, we can map the world and layer on the geo-located points of interest.

```{r pressure, echo=FALSE}
tm_shape(world) +
  tm_fill() +
  tm_shape (point_geo) +
  tm_bubbles(col = "red", size = 0.01) # Set the color and size of the geo-located points
```

## Making a better map
Also from the tmap package, the tmap_mode() function allows us to set the map on real geographical information, 
such as topography.

```{r pressure, echo=FALSE}
tmap_mode("view")

tmap_last()
```

## Calculate the distance

Then, we can use the st_distance() function from the sf package to find the distances between points.

```{r pressure, echo=FALSE}
dist <- st_distance(point_geo)
dist_km <- dist/1000
dist_km
```

## Save the map

Saving the map is easy using the tmap_save() function.

```{r pressure, echo=FALSE}
tmap_mode("view")

immigration_services <-  tmap_last()

tmap_save(immigration_services, "immigration_services.html")
```

## So let's make a map!

Some landmarks based on where students from the I2DS class are from:
India: ताजमहल (Taj Mahal)
Brazil: Cristo Redentor (Christ the Redeemer)
US: Empire State Building
UK: Big Ben
Italy: Colosseo (Colosseum)
Peru: Qurikancha (Golden Temple)
Panama: Vía Cincuentenario, Panamá, Panama (Old Cathedral at Panamá Viejo)
Mexico: Monumento a la Independencia (Independence Monument)
Netherlands: Erasmusbrug (Erasmus Bridge, Rotterdam)
Colombia: El Peñón de Guatapé (Rock of Guatape)
Germany: Fernsehturm, Berlin (TV Tower)
Bosnia: ??
Spain: ??

```{r pressure, echo=FALSE}
I2DS_students <- c("Friedrichstrasse 180, Berlin, Germany",
                   "ताजमहल",
               "Cristo Redentor",
               "Empire State Building",
               "Big Ben",
               "Colosseo",
               "Qurikancha",
               "Vía Cincuentenario, Panamá, Panama",
               "Monumento a la Independencia",
               "Erasmusbrug",
               "El Peñón de Guatapé",
               "Fernsehturm, Berlin",
               "Rock Creek Park, Washington, DC",
               "Parque do Ibirapuera, São Paulo") 

geocoded_I2DS <- geocode_OSM(I2DS_students)

geocoded_I2DS
```


## Setting up the geo-points

```{r pressure, echo=FALSE}
I2DS_points <- st_as_sf(geocoded_I2DS,
                      coords = c(x = "lon", y = "lat"),
                      crs = 4326) 

I2DS_points
```

Within the world data frame, there is a column (POP_EST) that is the population estimates for each country. I can therefore use this to add population data to the map using the tm_polygons() function.

```{r pressure, echo=FALSE}
tm_shape(world) +
  tm_polygons("POP_EST", id = "SOVEREINT", breaks = c(0, 1000000, 10000000, 50000000, 100000000, 200000000, 500000000, 1000000000, 1500000000)) +
  tm_fill() +
  tm_shape (I2DS_points) +
  tm_bubbles(col = "blue", size = 0.05) 

tmap_mode("view")
```

So, who is coming from the furtherest distance from Hertie??

```{r pressure, echo=FALSE}
I2DS_dist <- st_distance(I2DS_points)
I2DS_dist_km <- I2DS_dist/1000
I2DS_dist_km
```

```{r pressure, echo=FALSE}
I2DS_student_map <- tmap_last()

tmap_save(I2DS_student_map, "I2DS_student_map.html")
```

## FUN EXERCISES!! (WARNING: Learning to geocode and map in R will completely consume your life)

## Mapping

Make a country map in R.
You can download the shapefile (.shp) as a .zip file from the Internet. 
You can then convert this .shp file into a data frame that R can process.
Then, you can plot the data, which will for a map!

```{r pressure, echo=FALSE}
# Go to http://diva-gis.org/gdata and pick a country. 

country_file <- "" #Once you've picked a country, click "OK" then right click on the "Download button to "Copy Link", paste the hyperlink in the quotation marks. 

# Download .zip
download.file(country_file, destfile = "COUNTRY.zip")

# Unzip the file
unzip("COUNTRY.zip")

# Examine the file .zip file to find the shapefile
list.files()

# Save the shapefile as a dataframe
country <- sf::read_sf("") #Within the working directory, find the .shp file (look for ~adm1.shp for Diva GIS files) and use the read_sf() function to convert the spacial data into a dataframe. 

country

# Plot the map with population using ggplot
ggplot(country) + geom_sf() + theme_minimal()

#Or
qtm(country)

```

## Geocode

Think of addresses or landmarks in that country, which you would like to plot on the map

```{r pressure, echo=FALSE}
your_location <- c("") #Put the address or landmark in the quotations (remember to separate them with commas)

geocoded_your_location <- geocode_OSM(your_location) #geocode_OSM() from tmap package

geocoded_your_location
```

Now use the st_as_sf() function to create the geo-points
```{r pressure, echo=FALSE}
your_points <- st_as_sf(coords = c(x = "", y = ""), crs = 4326) 

your_points
```

Use the tm_shape() function to plot your selected country along with the point(s)
```{r pressure, echo=FALSE}
tm_shape(country) +
  tm_shape (your_points) +
  tm_bubbles(col = "blue", size = 0.05) 
```

Using the st_distance() function, calculate the distance between those points
```{r pressure, echo=FALSE}

```

## Bonus !!

Below is an example of how we can use data from the internet to make informative maps with geolocated points.
Let's plot the points discussed above on a map of Europe. 
But also overlaying this with spatial information on the uptake of asylum seekers per country in 2015.

```{r pressure, echo=FALSE}
#Import tha small data frame, which is taken from a table in Wikipedia (https://en.wikipedia.org/wiki/European_migrant_crisis) on asylum claims processed in different European countries
euro_ref <- read_excel("euro_ref.xlsx")

#Create a new column, which will allow us to join this data with the world data frame
euro_ref <- euro_ref %>% 
  dplyr::mutate(SOVEREIGNT_2 = state) %>%
  dplyr::select(3, 1, 2)

#Create a new column in the world data frame, which will allow us to join it with the euro_ref df
world_2 <- world %>%
  dplyr::mutate(SOVEREIGNT_2 = SOVEREIGNT)

#Join the data frames using the left_join() function and the new column created in each df
combined_ref <- left_join(x = world_2, 
                               y = euro_ref, 
                               by = "SOVEREIGNT_2")

#Filter so we only get a map of Europe
euro_ref_map <- combined_ref %>%
  filter(!is.na(combined_ref$state))

#Place the map and the points
tm_shape(euro_ref_map)+
  tm_polygons("num_ref", id = "state", palette = "Blues", breaks = c(0, 20000, 100000,  500000, 1000000, 1500000)) +
    tm_shape (point_geo) +
  tm_bubbles(col = "red", size = 0.001)

#Make sure to save !
european_ref <-  tmap_last()
tmap_save(european_ref, "european_ref.html")
```
