---
title: "Geocoding with R"
subtitle: "The SF package"
author: "Andrew Wells & Fernanda Candido Gomes"
institute: "Hertie School | [I2DS Tools for Data Science workshop](https://github.com/intro-to-data-science-21-workshop)"
date: "04th November 2021"
output: 
  html_document:
    toc: TRUE
    number_sections: FALSE
    highlight: tango
    theme: lumen
    toc_depth: 2
    toc_float:
      collapsed: false 
---

```{r setup, include=FALSE}
library(pacman)
pacman::p_load(sf, leaflets, readxl,
               dplyr, qtm, ggplot2, rvest, tidyverse, raster, tmap, tmaptools)
#raster is for mac

options(scipen = 999)
```

## Mapping

Let's make our first country map in R.
As Fernanda said, in the video, we can download the shapefile (.shp) as a .zip file from the Internet

```{r pressure, echo=FALSE}
# Go to http://diva-gis.org/gdata and pick a country. 

country_file <- " " #Once you've picked a country, click "OK" then right click on the "Download button to "Copy Link", paste the hyperlink in the parentheses. 

# Download .zip
download.file(country_file, destfile = "COUNTRY.zip")

# Unzip the file
unzip("COUNTRY.zip")

# Examine the file .zip file to find the shapefile
list.files()

# Save the shapefile as a dataframe
country <- sf::read_sf(" ") #Find the .shp file (look for ~adm1.shp for Diva GIS files) and use the read_sf() function to convert the spacial data into a dataframe. 

country

# Plot the map with population using ggplot
ggplot(country) + geom_sf() + theme_minimal()
```

## Finding coordinates from addresses

UNHCR Germany - Zimmerstrasse 80, Berlin, Germany
UNHCR Germany/Federal Office for Migrantion and Refugees - Frankenstrasse 210, Nurnberg, Germany
French Office of Immigration and Integration - 83 Rue de Patay, Paris, France
French Office of Immigration and Integration - 4 Rue Gustave Doré, Strasbourg, France
Belgium Immigration Office - Pachéco 44, Brussels, Belgium
Port of Antwerp Schengen Immigration Office - Kauwenstein 8, Antwerpen, Belgium
CARE - Switzerland - Chemin de Balexert 7, Geneva, Switzerland
CARE UK - 89 Albert Embankment, London, United Kingdom
Danish Refugee Council - Borgergade 10, København, Denmark
European Asylum Support Office - Valletta, Malta

```{r pressure, echo=FALSE}
addresses <- c("Zimmerstrasse 80, Berlin, Germany",
               "Frankenstrasse 210, Nurnberg, Germany",
               "83 Rue de Patay, Paris, France",
               "4 Rue Gustave Doré, Strasbourg, France",
               "Pachéco 44, Brussels, Belgium",
               "Kauwenstein 8, Antwerpen, Belgium",
               "Chemin de Balexert 7, Geneva, Switzerland",
               "89 Albert Embankment, London, United Kingdom",
               "Borgergade 10, København, Denmark",
               "Valletta, Malta") 

geocoded_addresses <- geocode_OSM(addresses)

geocoded_addresses
```

## Setting up the geo-points

Use the st_as_sf() function

```{r pressure, echo=FALSE}
point_geo <- st_as_sf(geocoded_addresses,
                      coords = c(x = "lon", y = "lat"),
                      crs = 4326)
point_geo
```

## Downloading a map of the world
diva-gis.org is great for individual countries, but does not have a files for more than one country. For this, let's go to Natural Earth (naturalearthdata.com). 

Click on the 1:50m Cultural Vectors. Right click on the "Download countries" button and copy the link. 

```{r pressure, echo=FALSE}

world_file <- "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/50m/cultural/ne_50m_admin_0_countries.zip" 

# Download .zip
download.file(world_file, destfile = "WORLD.zip")

# Unzip the file
unzip("WORLD.zip")

# Examine the file .zip file to find the shapefile
list.files()

# Save the shapefile as a dataframe
world <- sf::read_sf("ne_50m_admin_0_countries.shp") #Look for the .shp
world
```


## Make a quick map of the world

```{r pressure, echo=FALSE}
ggplot(world) + geom_sf() + theme_minimal()
```


## Joining dataframes

Use the st_join() function

```{r pressure, echo=FALSE}
st_join(point_geo, world, join = st_within)
```

## Map it all together!


```{r pressure, echo=FALSE}
tm_shape(world) +
  tm_fill() +
  tm_shape (point_geo) +
  tm_bubbles(col = "red", size = 0.01)
```

## Making a better map


```{r pressure, echo=FALSE}
tmap_mode("view")

tmap_last()
```

## Calculate the distance

Use the st_distance() function to find the distances between points.

```{r pressure, echo=FALSE}
dist <- st_distance(point_geo)
dist_km <- dist/1000
dist_km
```


## Save the map

Saving the map is easy using the tmap_save() function.

```{r pressure, echo=FALSE}
tmap_mode("view")

immigration_services <-  tmap_last()

tmap_save(immigration_services, "immigration_services.html")
```

## Bonus !!

Let's plot the points on a map of Europe. 
But also overlaying this with spatial information on the uptake of asylum seekers per country in 2015.

```{r pressure, echo=FALSE}
euro_ref <- read_excel("~/Downloads/euro_ref.xlsx")

euro_ref <- euro_ref %>% 
  dplyr::mutate(SOVEREIGNT_2 = state) %>%
  dplyr::select(4, 1, 2, 3)

world_2 <- world %>%
  dplyr::mutate(SOVEREIGNT_2 = SOVEREIGNT)

combined_ref <- left_join(x = world_2, 
                               y = euro_ref, 
                               by = "SOVEREIGNT")

euro_ref_map <- combined_ref %>%
  filter(!is.na(combined_ref$state))
  
tm_shape(euro_ref_map)+
  tm_polygons("num_ref", id = "state", palette = "Greens") +
    tm_shape (point_geo) +
  tm_bubbles(col = "red", size = 0.001)

# Make sure to save !
european_ref <-  tmap_last()
tmap_save(european_ref, "european_ref.html")
```